\begin{hcarentry}{binary}
\label{binary}
\report{Lennart Kolmodin}
\status{active}
\participants{Duncan Coutts, Don Stewart, Binary Strike Team}
\entry{unchanged, pinged}% done 2007-12-22
\makeheader

The Binary Strike Team is pleased to announce the release of a new,
pure, efficient binary serialisation library.

The `binary' package provides efficient serialisation of Haskell values
to and from lazy ByteStrings. ByteStrings constructed this way may then
be written to disk, written to the network, or further processed (e.g.
stored in memory directly, or compressed in memory with zlib or bzlib).

The binary library has been heavily tuned for performance, particularly for
writing speed. Throughput of up to 160M/s has been achieved in practice, and
in general speed is on par or better than NewBinary, with the advantage of a
pure interface. Efforts are underway to improve performance still further.
Plans are also taking shape for a parser combinator library on top of
binary, for bit parsing and foreign structure parsing (e.g. network
protocols).

Data.Derive~\cref{derive} has support for automatically generating Binary
instances, allowing to read and write your data structures with little fuzz.

Binary was developed by a team of 8 during the Haskell Hackathon, and since
then has in total 15 people contributed code and many more given feedback
and cheerleading on \verb|#haskell|.

The underlying code is currently being rewritten to give even better
performance -- both reading and writing -- still exposing the same API.

The package is is available through Hackage~\cref{hackagedb}.

\FurtherReading
\begin{compactitem}
\item Homepage

  \url{http://www.cse.unsw.edu.au/~dons/binary.html}
\item Hackage

  \url{http://hackage.haskell.org/cgi-bin/hackage-scripts/package/binary}
\item Development version

  \texttt{darcs get --partial}

  \url{http://darcs.haskell.org/binary}
\end{compactitem}
\end{hcarentry}
